Waveform and Passive Beamforming Design for Intelligent Reflecting Surface-Aided Wireless Information and Power Transfer

### 2020-09-21 main.tex
:pencil: R-E region and splitting ratio vs noise

		To maximize non-concave expression \ref{eq:z_irs}, we propose a Successive Convex Approximation (SCA)-based algorithm that approximate the second-order terms by first-order Taylor expansion. Based on the variables optimized at iteration $i - 1$, the local approximation at iteration $i$ suggests \cite{Adali2010}
		\begin{align}
			(t_{I,0}^{(i)})^2
			& \ge 2 t_{I,0}^{(i)}t_{I,0}^{(i-1)} - (t_{I,0}^{(i-1)})^2\label{eq:taylor_1}\\
			t_{P,n}^{(i)} (t_{P,n}^{(i)})^*
			& \ge 2 \Re\left\{t_{P,n}^{(i)} (t_{P,n}^{(i-1)})^*\right\} - t_{P,n}^{(i-1)} (t_{P,n}^{(i-1)})^*\label{eq:taylor_2}\\
			t_{I,0}^{(i)} t_{P,0}^{(i)}
			& = \frac{1}{4}(t_{I,0}^{(i)} + t_{P,0}^{(i)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\nonumber\\
			& \ge \frac{1}{2}(t_{I,0}^{(i)} + t_{P,0}^{(i)})(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})\nonumber\\
			& \quad - \frac{1}{4}(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\label{eq:taylor_3}
		\end{align}
		which provide lower bounds to the corresponding terms in \ref{eq:z_irs}. At iteration $i$, the approximated objection function $\tilde{z}(\boldsymbol{\Phi}^{(i)})$ is detailed in \ref{eq:z_irs_approx}.
		\begin{figure*}[b]
			\hrule
			\begin{equation}\label{eq:z_irs_approx}
				\begin{split}
					\tilde{z}(\boldsymbol{\Phi}^{(i)})
					& = \frac{1}{2}{\beta_2}{\rho}(t_{I,0}^{(i)}+t_{P,0}^{(i)})\\
					& \quad + \frac{3}{8}{\beta_4}{\rho^2} \left(4 (t_{I,0}^{(i)})(t_{I,0}^{(i-1)}) - 2 (t_{I,0}^{(i-1)})^2 + \sum_{n=-N+1}^{N-1}{2 \Re\left\{t_{P,n}^{(i)} (t_{P,n}^{(i-1)})^*\right\} - t_{P,n}^{(i-1)} (t_{P,n}^{(i-1)})^*}\right)\\
					& \quad + \frac{3}{2}{\beta_4}{\rho^2} \left(\frac{1}{2}(t_{I,0}^{(i)} + t_{P,0}^{(i)})(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)}) - \frac{1}{4}(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\right)
				\end{split}
			\end{equation}
		\end{figure*}
		Hence, problem \ref{op:original} is transformed to
		\begin{maxi!}
			{\boldsymbol{\boldsymbol{\Phi}}}{\tilde{z}(\boldsymbol{\Phi})}{\label{op:irs}}{\label{ob:irs}}
			\addConstraint{R(\boldsymbol{\Phi}) \ge \bar{R}}\label{co:irs_rate}
			\addConstraint{\boldsymbol{\Phi}_{l,l}=1, \quad l=1,\dots,L+1}\label{co:irs_modulus}
			\addConstraint{\boldsymbol{\Phi}\succeq{0}}
			\addConstraint{\mathrm{rank}(\boldsymbol{\Phi})=1\label{co:irs_rank}}
		\end{maxi!}

		Problem \ref{op:irs} is not a standard Semidefinite Programming (SDP) due to the rate constraint \ref{co:irs_rate}. Therefore, even if we relax the rank constraint \ref{co:irs_rank} to formulate a convex problem, there is no guarantee that the optimal rank-1 solution $\bar{\boldsymbol{\phi}}^{\star}$ extracted from $\boldsymbol{\Phi}^{\star}$ is a stationary point of the original problem \ref{op:original}. In Section [TODO], we numerically show that $\boldsymbol{\Phi}^{\star}$ is rank-1 for all tested channel realizations so that the performance loss is insignificant.

@@

\begin{section}{Performance Evaluations}
	\begin{figure}
		\centering
		\def\svgwidth{\columnwidth}
		\import{assets/}{layout.pdf_tex}
		\caption{System layout}
		\label{fi:layout}
	\end{figure}
	To evaluate the performance of the proposed IRS-aided SWIPT system, we characterize and compare the average R-E regions under typical setups. Consider a large open space WiFi-like environment at a center frequency of 5.18 \si{\GHz} with reference bandwidth $B=1$ \si{\MHz}. As shown in Fig. \ref{fi:layout}, we assume the IRS moves along a horizontal line parallel to the AP-user path. Define the projection of the AP-IRS path onto the AP-user path as $d_H$ and let the distance from the IRS to the AP-user path be $d_V$. Denote $d_D$, $d_I$, $d_R$ as the length of direct, incident, reflective paths, such that $d_I=\sqrt{d_H^2+d_V^2}$, $d_R=\sqrt{(d_D-d_H)^2+d_V^2}$. We fix $d_D=15$ \si{\meter}, $d_V=2$ \si{\meter} and choose a reference $d_H=2$ \si{\meter}. For all channels, the path loss and fading parameters are obtained from IEEE TGn channel model D \cite{Erceg2004}, and we regard all channels as NLoS unless otherwise mentioned. Reference path loss is set to $L_0=-35$ \si{\dB} at $d_0=1$ \si{\meter}, and taps are modelled as i.i.d. random variables with normalized total power. We choose number of subbands $N=16$ as reference and assume no spatial correlation across all arrays. Rectenna parameters are taken as $k_2=0.0034$, $k_4=0.3829$, $R_{\text{ant}}=50$ \si{\ohm}. The average Effective Isotropic Radiated Power (EIRP) is fixed to $-36$ \si{\dBm} while the reference average noise power is $\sigma_n=-40$ \si{\dBm} at all subbands. We also assume 0 \si{\dBi} IRS element gain and 2 \si{\dBi} receive antenna gain. For the algorithm, the tolerance is $\epsilon=10^{-8}$, the number of candidates in the Gaussian randomization method is $Q=10^{3}$, and the R-E region is averaged over 200 channel realizations. In the R-E boundary, the leftmost point corresponds to WPT ($\rho=1$) where power can be allocated simultaneously to modulated and unmodulated waveform to maximize the average output DC current. On the other hand, the rightmost point corresponds to WIT ($\rho=0$) where the solution coincides with water-filling algorithm allocating all power to modulated waveform only. For a fair comparison, the $x$-axis of the plots has been normalized to per-subband rate.

@@

	% N increase -> more to balance, tradeoff between subbands
	% FF channel -> closed form optimal IRS; FS channel -> tradeoff between subbands

	\begin{figure}
		\centering
		\begin{adjustbox}{width=\linewidth}
			\input{assets/re_noise.tex}
		\end{adjustbox}
		\caption{Average R-E region versus the average noise power}
		\label{fi:re_noise}
	\end{figure}

	The influence of the average noise power on the average R-E region is investigated in Fig. \ref{fi:re_noise}. It is noted that for a large number of subbands ($N=16$ in this case), the R-E region is approximately concave for a high noise level and approximately convex for a low noise level. Moreover, there exists a

	% This is because

	As a result, TS is preferred at low SNR while PS is preferred at high SNR.

	% sections -> waterfilling: \rho only vs joint

	% irs, low noise power, high received power  L+->high snr, rho vs other L, M: snr and nonlinearity
	% frequency selective less multipath loss less with FF IRS
	% Fix subband rate

@@

	\begin{figure}
		\centering
		\begin{adjustbox}{width=\linewidth}
			\input{assets/re_los.tex}
		\end{adjustbox}
		\caption{Average R-E region for IRS-aided NLoS and LoS channels}
		\label{fi:re_los}
	\end{figure}

@@

	% TODO: FF channels in IRS plot
\end{section}

>>>>

		To maximize non-concave expression \ref{eq:z_irs}, we propose a Successive Convex Approximation (SCA)-based algorithm that approximate the second-order terms by first-order Taylor expansion. Based on the variables optimized at iteration $i - 1$, the local approximation at iteration $i$ suggests \cite{Adali2010}
		\begin{align}
			(t_{I,0}^{(i)})^2
			& \ge 2 t_{I,0}^{(i)}t_{I,0}^{(i-1)} - (t_{I,0}^{(i-1)})^2\label{eq:taylor_1}\\
			t_{P,n}^{(i)} (t_{P,n}^{(i)})^*
			& \ge 2 \Re\left\{t_{P,n}^{(i)} (t_{P,n}^{(i-1)})^*\right\} - t_{P,n}^{(i-1)} (t_{P,n}^{(i-1)})^*\label{eq:taylor_2}\\
			t_{I,0}^{(i)} t_{P,0}^{(i)}
			& = \frac{1}{4}(t_{I,0}^{(i)} + t_{P,0}^{(i)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\nonumber\\
			& \ge \frac{1}{2}(t_{I,0}^{(i)} + t_{P,0}^{(i)})(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})\nonumber\\
			& \quad - \frac{1}{4}(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\label{eq:taylor_3}
		\end{align}
		which provide lower bounds to the corresponding terms in \ref{eq:z_irs}. At iteration $i$, the approximated objection function $\tilde{z}(\boldsymbol{\Phi}^{(i)})$ is detailed in \ref{eq:z_irs_approx}.
		\begin{figure*}[b]
			\hrule
			\begin{equation}\label{eq:z_irs_approx}
				\begin{split}
					\tilde{z}(\boldsymbol{\Phi}^{(i)})
					& = \frac{1}{2}{\beta_2}{\rho}(t_{I,0}^{(i)}+t_{P,0}^{(i)})\\
					& \quad + \frac{3}{8}{\beta_4}{\rho^2} \left(4 (t_{I,0}^{(i)})(t_{I,0}^{(i-1)}) - 2 (t_{I,0}^{(i-1)})^2 + \sum_{n=-N+1}^{N-1}{2 \Re\left\{t_{P,n}^{(i)} (t_{P,n}^{(i-1)})^*\right\} - t_{P,n}^{(i-1)} (t_{P,n}^{(i-1)})^*}\right)\\
					& \quad + \frac{3}{2}{\beta_4}{\rho^2} \left(\frac{1}{2}(t_{I,0}^{(i)} + t_{P,0}^{(i)})(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)}) - \frac{1}{4}(t_{I,0}^{(i-1)} + t_{P,0}^{(i-1)})^2 - \frac{1}{4}(t_{I,0}^{(i)} - t_{P,0}^{(i)})^2\right)
				\end{split}
			\end{equation}
		\end{figure*}
		Hence, problem \ref{op:original} is transformed to
		\begin{maxi!}
			{\boldsymbol{\Phi}}{\tilde{z}(\boldsymbol{\Phi})}{\label{op:irs}}{\label{ob:irs}}
			\addConstraint{R(\boldsymbol{\Phi}) \ge \bar{R}}\label{co:irs_rate}
			\addConstraint{\boldsymbol{\Phi}_{l,l}=1, \quad l=1,\dots,L+1}\label{co:irs_modulus}
			\addConstraint{\boldsymbol{\Phi}\succeq{0}}
			\addConstraint{\mathrm{rank}(\boldsymbol{\Phi})=1\label{co:irs_rank}}
		\end{maxi!}

		Problem \ref{op:irs} is not a standard Semidefinite Programming (SDP) due to the rate constraint \ref{co:irs_rate}. Therefore, even if we relax the rank constraint \ref{co:irs_rank} to formulate a convex problem, there is no guarantee that the optimal rank-1 solution $\bar{\boldsymbol{\phi}}^{\star}$ extracted from $\boldsymbol{\Phi}^{\star}$ is a stationary point of the original problem \ref{op:original}. In Section \ref{se:performance_evaluation}, we numerically show that $\boldsymbol{\Phi}^{\star}$ is rank-1 for all tested channel realizations so that the performance loss is insignificant.

@@

\begin{section}{Performance Evaluations}\label{se:performance_evaluation}
	\begin{figure}
		\centering
		\def\svgwidth{\columnwidth}
		\import{assets/}{layout.pdf_tex}
		\caption{System layout}
		\label{fi:layout}
	\end{figure}
	To evaluate the performance of the proposed IRS-aided SWIPT system, we characterize and compare the average R-E regions under typical setups. Consider a large open space WiFi-like environment at a center frequency of 5.18 \si{\GHz} with reference bandwidth $B=1$ \si{\MHz}. As shown in Fig. \ref{fi:layout}, we assume the IRS moves along a horizontal line parallel to the AP-user path. Define the projection of the AP-IRS path onto the AP-user path as $d_H$ and let the distance from the IRS to the AP-user path be $d_V$. Denote $d_D$, $d_I$, $d_R$ as the length of direct, incident, reflective paths, such that $d_I=\sqrt{d_H^2+d_V^2}$, $d_R=\sqrt{(d_D-d_H)^2+d_V^2}$. We fix $d_D=15$ \si{\meter}, $d_V=2$ \si{\meter} and choose a reference $d_H=2$ \si{\meter}. For all channels, the path loss and fading parameters are obtained from IEEE TGn channel model D \cite{Erceg2004}, and we regard all channels as NLoS unless otherwise mentioned. Reference path loss is set to $L_0=-35$ \si{\dB} at $d_0=1$ \si{\meter}, and taps are modelled as i.i.d. random variables with normalized total power. We choose number of subbands $N=16$ as reference and assume no spatial correlation across all arrays. Rectenna parameters are taken as $k_2=0.0034$, $k_4=0.3829$, $R_{\text{ant}}=50$ \si{\ohm}. The average Effective Isotropic Radiated Power (EIRP) is fixed to $-36$ \si{\dBm} while the reference average noise power is $\sigma_n=-40$ \si{\dBm} at all subbands. We also assume 0 \si{\dBi} IRS element gain and 2 \si{\dBi} receive antenna gain. For the algorithm, the tolerance is $\epsilon=10^{-8}$, the number of candidates in the Gaussian randomization method is $Q=10^{3}$, and the R-E region is averaged over 200 channel realizations. In the R-E boundary, the leftmost point corresponds to WPT ($\rho=1$) where power can be allocated simultaneously to modulated and unmodulated waveform to maximize the average output DC current. On the other hand, the rightmost point corresponds to WIT ($\rho=0$) where the solution coincides with water-filling algorithm allocating all power to modulated waveform only. For a fair comparison, the $x$-axis of the plots has been normalized to per-subband rate.

	We first evaluate the performance of Algorithm \ref{al:irs} under SDR. It is demonstrated that $\boldsymbol{\Theta}^{\star}$ is rank-1 for all tested channel realizations with different $M$, $N$ and $L$. Therefore, $\boldsymbol{\theta}^{\star}$ can be directly obtained through eigenvalue decomposition and we claim Algorithm \ref{al:irs} converges to stationary points of problem \ref{op:irs} without performance loss.

@@

	\begin{figure}
		\centering
		\begin{adjustbox}{width=\linewidth}
			\input{assets/re_noise.tex}
		\end{adjustbox}
		\caption{Average R-E region and splitting ratio versus the average noise power}
		\label{fi:re_noise}
	\end{figure}

	The influence of the average noise power on the average R-E region is investigated in Fig. \ref{fi:re_noise}. We \textit{first} note that for a large number of subbands ($N=16$), the R-E region is approximately concave for a high noise level and approximately convex for a low noise level. Hence, TS is preferred at low SNR while PS is preferred at high SNR. This is because at a low SNR, the capacity-achieving Water-Filling (WF) algorithm tends to allocate more power to few strongest subbands. As the rate constraint $\bar{R}$ decreases, more subbands are activated such that the rectifier nonlinearity further boosts the harvested energy by introducing more balanced terms to the output DC current. \textit{Second}, there exists a turning point in the R-E region especially for a small noise ($\sigma_n \le -40$ \si{\dBm}). The reason is that when $\bar{R}$ decreases slightly from the capacity, the algorithm mainly adjusts the splitting ratio $\rho$ rather than put more weight on the multisine waveform, since a small amplitude could be inefficient for energy maximization. On the other hand, as $\bar{R}$ further reduces, a modulated waveform with a very large $\rho$ could be outperformed by a composed waveform with a smaller $\rho$, due to advantage the of multisine. The result highlights the necessity of joint optimization of waveform and splitting ratio.

@@

@@

	% Same reflection coefficient, When b small, even if n large, behaves as if ff
	% FF channel -> closed form optimal IRS; FS channel -> tradeoff between subbands
	% \textit{Third}, although each IRS element has the same reflection coefficient for $N$ subbands, increasing $N$ does not necessarily mean
	% frequency selective less multipath loss less with FF IRS

\end{section}
